---
title: "Lab 04 - La Quinta is Spanish for next to Denny's, Pt. 2"
author: "Kyle Knaut"
date: "January 29th, 2024"
output: github_document
---

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
library(dsbox) 
```

```{r load-data, message = FALSE}
states <- read_csv("data/states.csv")
```

### Exercise 1
After filtering the set, there are three Denny's locations in Alaska

```{r alaska_filter}
dn_ak <- dennys %>%
  filter(state == "AK")
nrow(dn_ak)

```
### Exercise 2
There are only two La Quinta's in Alaska
```{r alaska_filter_lq}
lq_ak <- laquinta %>%
  filter(state == "AK")
nrow(lq_ak)
```
### Exercise 3
There are 6 possible pairings of Denny's and Laquintas in Alaska that we will calculate distances for. 

### Exercise 4
There are 6 observations in this new dataframe, consistent with what I would expect. The variables inlcude address, city, state, zipcode, longitude, and lattitude. I'm slightly confused because there are actually eleven variables being observed, where there are six observations of addressed from the Denny's dataset, and six observations from the Laquinta dataset, even though both of these have two and three respectively, but what I hope is happening is that each observation is a pair of locations.
``` {r joining_datasets}
dn_lq_ak <- full_join(dn_ak, lq_ak, by = "state")
dn_lq_ak


```
### Exercise 5

In order to add a new variable to this dataframe without altering the other data, I would use the mutate function from tidyverse.

### Exercise 6
```{r haversine_fct}
haversine <- function(long1, lat1, long2, lat2, round = 3) {
  # convert to radians
  long1 <- long1 * pi / 180
  lat1 <- lat1 * pi / 180
  long2 <- long2 * pi / 180
  lat2 <- lat2 * pi / 180

  R <- 6371 # Earth mean radius in km

  a <- sin((lat2 - lat1) / 2)^2 + cos(lat1) * cos(lat2) * sin((long2 - long1) / 2)^2
  d <- R * 2 * asin(sqrt(a))

  return(round(d, round)) } 
#Stolen from SMG :)
```

Adding distance variable, ask SMG what unit of measure haversine spits out.
```{r distance}
distance_ak <- dn_lq_ak %>%
  mutate(distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y, round = 3))

distance_ak

```

### Exercise 7
``` {r minimun_distance}
dn_lq_ak_mindist <- distance_ak %>%
  group_by(address.x) %>%
  summarize(closest = min(distance))

dn_lq_ak_mindist

```

### Exercise 8
``` {r visualising_establishment_distances}
ggplot(dn_lq_ak) + geom_point(aes(x = longitude.x, y = latitude.x, col = "red")) + geom_point(aes(x = longitude.y, y = latitude.y, col = "black")) + labs(x = "Longitude", y = "Latitude", color = "Establishment", title = "Distribution of Dennys / Laquintas in Alaska") + theme_minimal() 
  
  #legend(1, 95, legend=c("Denny's", "La Quinta"), col=c("red", "black"))

ggplot(dn_lq_ak_mindist, mapping = aes(x = address.x, y = closest, color = address.x)) + geom_boxplot() + labs(x = "Denny's address", y = "Distance to nearest La Quitna", title = "Distance from Denny's to Laquintas in Alaska") + theme_minimal() + guides(color = FALSE)

#ggplot() + layer(data = dn_lq_ak, mapping = aes(x = latitude.x, y = longitude.x, groupName =="Dennys", groupColors=c('#999999')), geom_point(), stat = "address", position = "address")

#dn_lq_ak %>%
  #ggplot(mapping = aes(x = longitude.x, y = latitude.x, color = "#999")) + geom_point() + labs(x = "Longitude", y = "Latitude", color = "Establishment", title = "Distribution of Dennys / Laquintas in Alaska") 
#ggplot(dn_lq_ak, mapping = aes(x = longitude.y, y = latitude.y, color = "#111")) + geom_point()
```

Looking at these visualizations, it seems like Denny's and La Quinta do tend to be quite close to eachother, with none being further than five (units) apart, and one pair being only 2 (units) apart. It appears that Denny's and La Quinta are both located in only two cities, Fairbanks, and Anchorage, and none of these locations are very far apart because of this. 

### Exercise 9
``` {r NC_analysis}
#Filtering for NC
dn_nc <- dennys %>%
  filter(state == "NC")
nrow(dn_nc)
# This time, there are far more Denny's locations, with 28 observations across the state

lq_nc <- laquinta %>%
  filter(state == "NC")
nrow(lq_nc)
#There are also more La Quinta's, although not in the same proportions, with only 12 locations

#joining the data sets
dn_lq_nc <- full_join(dn_nc, lq_nc, by = "state")
dn_lq_nc

distance_nc <- dn_lq_nc %>%
  mutate(distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y, round = 3))

distance_nc
```

### Exercise 10
``` {r Texas_analysis}

```

### Exercise 11
``` {r Colorado_analysis}

```

### Exercise 12
Mitch Hedberg's joke is most true in ______
